---
title: "Our Data Analysis"
output:
  html_document: 
    toc: yes
    toc_float: yes
    keep_md: yes
---

These are the steps we took to analyze our data.  



----
## **Load the libraries**

```{r}
library(tidyverse)
library(janitor)
library(here)
library(naniar)
library(skimr)
```

<br>
<br>

## **Cleaning Up the Data**

### Load the data frame


```{r}
covid <- readr::read_csv("data/WHO-COVID-19-global-data.csv")
```


### Clean the data up
```{r}
covid<-janitor::clean_names(covid)
```

```{r}
covid$date_reported <- lubridate::mdy(covid$date_reported)
```

```{r}
covid$date_reported <- as.Date(covid$date_reported)
covid$cumulative_cases <- as.numeric(as.character(covid$cumulative_cases))
```

<br>
<br>

## **Preliminary analysis**
```{r}
skim(covid)
```
<font size="3" color="lightblue4">The NA's in this data are represented by "NA" and "######"</font>

<br>
<br>

## **Cumulative Cases**
### Which countries had the most cumulative cases?
<font size="3" color="blue">This shows the data from the most recent data collection date: 02-11-2021..</font>
```{r}
covid %>% 
  filter(date_reported == "2021-02-11") %>% 
  arrange(desc(cumulative_cases))
```
<font size="3" color="blue">The countries that have the most cumulative cases in the world are: the United States, India, Brazil, the Russian Federation, and the United Kingdom.</font>


<br>
<br>


### Cumulative cases in US over time
```{r}
covid%>%
  filter(country_code == "US")%>%
  ggplot(aes(x = date_reported, y = cumulative_cases))+
  geom_line(aes(group = 1))+
  scale_x_date(date_labels = "%Y-%m-%d")+
  scale_x_date(date_breaks = "1 month")+
  scale_y_continuous(breaks = seq(0, 30000000, by = 2500000), limits=c(0, 30000000))+
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  scale_color_brewer(palette = "Set2")+
  labs(title="Cumulative Cases over Time in the US", x="Date Reported", y="Cumulative Cases")+
  theme(plot.title = element_text(size = rel(1.26), hjust = 0.5))
```

<font size="3" color="blue">The United States had the most cumulative cases, so we decided to plot the number of cases in the US from the beginning of the data collection (01-01-2020) to the most recent data we have (02-11-2021).</font>

<br>
<br>

### Cumulative cases in the world over time
```{r}
covid%>%
  group_by(country)%>%
  filter(country_code == "US"| country_code == "IN"| country_code == "BR"| country_code == "RU"| country_code == "GB")%>%
  ggplot(aes(x = date_reported, y = cumulative_cases, group = country, color = country))+
  geom_line()+
  scale_x_date(date_labels = "%Y-%m-%d")+
  scale_x_date(date_breaks = "1 month")+
  scale_y_continuous(breaks = seq(0, 30000000, by = 2500000), limits=c(0, 30000000))+
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  scale_color_brewer(palette = "Set2")+
  labs(title="Cumulative Cases over Time", x="Date Reported", y="Cumulative Cases")+
  theme(plot.title = element_text(size = rel(1.26), hjust = 0.5))
```

<font size="3" color="blue">This plot shows us how the cumulative cases have increased for the top 5 countries with the most cumulative cases.</font>

<br>
<br>

### Which country's cumulative cases are growing the fastest?

<font size="3" color="blue">We did this by subtracting the amount of cases on 11-11-2020 from the amount of cases on 2-11-2021. This data is for the last 3 months.</font>
```{r}
covid%>%
  group_by(country)%>%
  filter(cumulative_cases > 1000)%>%
  filter(date_reported == "2020-11-11" | date_reported == "2021-2-11")%>%
  arrange(date_reported)%>%
  mutate(inc_rate =(cumulative_cases - lag(cumulative_cases)) / lag(cumulative_cases))%>%
  arrange(desc(inc_rate))
```

<font size="3" color="blue">In the last 3 months, the countries with the fastest growth of cumulative cases are: Uruguay, Latvia, Estonia, Lithuania, and Thailand.</font>

<br>
<br>

### Comparing the cumulative cases through faceting
```{r}
covid2 <- covid%>%
  filter(country == "Uruguay"| country == "Latvia" | country == "Estonia" | country == "Lithuania" | country == "Thailand")%>%
  filter(date_reported > "2020-11-11" , date_reported < "2021-2-11")%>%
  ggplot(aes(x = date_reported, y = cumulative_cases))+
  geom_line()+
  scale_x_date(date_labels = "%Y-%m-%d")+
  theme(axis.text.x=element_text(angle=45, hjust=1))+
  scale_fill_brewer(palette = "Set2")+
  labs(title = "Cumulative Cases in the Last 3 Months", x = "Date Reported", y = "Cumulative Cases")+
  theme(plot.title = element_text(size = rel(1.26), hjust = 0.5))
```
```{r}
covid2 + facet_wrap(~country)
```

<font size="3" color="blue">This plot shows us how the cumulative cases have increased for the countries that have increased in cumulative cases the fastest. We analyzed the last 3 months.</font>

<br>
<br>

## **Cumulative Deaths**
### Which countries had the most deaths?
```{r}
covid %>% 
  filter(date_reported== "2021-02-11") %>% 
  arrange(desc(cumulative_deaths))
```
<font size="3" color="blue">The countries with the highest cumulative deaths are: the United States, Brazil, Mexico, India, and the United Kingdom.</font>

<br>
<br>

### Cumulative deaths over time
```{r}
covid%>%
  group_by(country)%>%
  filter(country_code == "US"| country_code == "BR"| country_code == "MX"| country_code == "IN"| country_code == "GB")%>%
  ggplot(aes(x = date_reported, y = cumulative_deaths, group = country, color = country))+
  geom_line()+
  scale_x_date(date_labels = "%Y-%m-%d")+
  scale_x_date(date_breaks = "1 month")+
  scale_y_continuous(breaks = seq(0, 300000, by = 25000), limits=c(0, 300000))+
  theme(axis.text.x=element_text(angle=45, hjust=1))+
   scale_color_brewer(palette = "Set2")+
  labs(title="Cumulative Deaths over Time", x="Date Reported", y="Cumulative Deaths")+
  theme(plot.title = element_text(size = rel(1.26), hjust = 0.5))
```

<br>
<br>

## **New Cases**
### Countries with the highest amount of new cases
<font size="3" color="blue">This is data that was reported on 02-11-2021.</font> 
```{r}
covid%>%
  select(date_reported, country, new_cases)%>%
  filter(date_reported=="2021-02-11")%>%
arrange(desc(new_cases))
```
<font size="3" color="blue">The top 5 countries are : the United States, Brazil, France, the Russian Federation, and the United Kingdom.</font> 

<br>
<br>

### How New Cases on 02-11-2021 Compare in the Top 5 Countries
```{r}
covid%>%
  filter(country=="United States of America"| country=="Brazil"| country=="Russian Federation" |country=="The United Kingdom"| country=="France")%>%
  filter(date_reported=="2021-02-11")%>%
  ggplot(aes(x=date_reported, y=new_cases, group=country, fill=country))+
  geom_col(position="dodge")+scale_fill_brewer(palette = "Set2")+
  labs(title = "Countries with highest reported new cases",
       x="Date",
       y="New Cases")+
  theme(plot.title = element_text(size = rel(1.26), hjust = 0.5))
```

<font size="3" color="blue">This plot shows the number of new cases reported on February 11, 2021 across the countries that reported the most cases.</font> 

<br>
<br>

### Comparing New Cases in different WHO regions
<font size="3" color="blue">TThe WHO divides the world into 6 regions: African Region (AFRO), Region of the Americas (AMRO), South-East Asian Region (SEAR), the European Region (EURO), the Eastern Mediterranean Region (EMRO), and the Western Pacific Region (WPRO).</font>

```{r}
options(scipen=999) 
covid%>%
  group_by(who_region)%>%
  filter(who_region!="Other")%>%
  select(date_reported, new_cases, who_region)%>%
  ggplot(aes(x=who_region, y=new_cases, fill=who_region))+
  geom_col()+scale_fill_brewer(palette = "Set2")+
  labs(title = "Reported cases by WHO region as of 2/11/2021 ",
       x="WHO Region",
       y="New Cases")+
  theme(plot.title = element_text(size = rel(1.26), hjust = 0.5))
```

<br>
<br>

### Countries with the highest amount of new deaths

```{r}
covid%>%
  select(date_reported, country, new_deaths)%>%
  filter(date_reported=="2021-02-11")%>%
arrange(desc(new_deaths))
```

<font size="3" color="blue">The top 5 countries are : the United States, Mexico, Brazil, the United Kingdom, and Germany.</font> 

<br>
<br>

## **New Deaths**
### How New Deaths on 02-11-2021 Compare in the Top 5 Countries

```{r}
options(scipen=999) 
covid%>%
  filter(country=="United States of America"| country=="Mexico"| country=="Brazil" |country=="The United Kingdom"| country=="Germany")%>%
  filter(date_reported=="2021-02-11")%>%
  ggplot(aes(x=date_reported, y=new_deaths, group=country, fill=country))+
  geom_col(position="dodge")+scale_fill_brewer(palette = "Set2")+
  labs(title = "Countries with highest reported new deaths",
       x="Date",
       y="New Deaths")+
  theme(plot.title = element_text(size = rel(1.26), hjust = 0.5))
```

<font size="3" color="blue">This plot shows the number of new deaths reported on February 11, 2021 across the countries that reported the most deaths</font> 

<br>
<br>

### Comparing New Deaths by different WHO regions

```{r}
options(scipen=999) 
covid%>%
  group_by(who_region)%>%
  filter(who_region!="Other")%>%
  select(date_reported, new_deaths, who_region)%>%
  ggplot(aes(x=who_region, y=new_deaths, fill=who_region))+
  geom_col()+scale_fill_brewer(palette = "Set2")+
  labs(title = "Reported deaths by WHO region as of February 11 2021",
       x="WHO region",
       y="New Deaths")+
  theme(plot.title = element_text(size = rel(1.26), hjust = 0.5))
```


<br>
<br>
<br>
[Go Home](https://v-srinidhi.github.io/group8-finalproject/)
